
# Using Containers

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g2effc5b673e_0_833")
```

Let's actually dive into to using containers. You can follow along with this activity with either Docker or Podman. We recommend using Docker but if for admin reasons you are unable to use Docker, Podman's commands work nearly identically to Docker.

### A note about "window juggling"

When learning another software skill like containers, it sometimes means you'll have to keep track of more windows. We call this "Window Juggling". It's an overlooked skill but something you'll become more comfortable with.

**IMPORTANT**: Use the checkboxes to follow along with each step in the activities! This is to help you keep track because steps have to be done and in order!

## {.tabset}

### Docker
#### Set up
<input type="checkbox"> Go here to [install Docker](https://www.docker.com/get-started), following the instructions for your particular operating system.    
<input type="checkbox"> If you don't have a Docker account create an account when prompted, or [go here](https://hub.docker.com/).    
<input type="checkbox"> After you install Docker, start up Docker desktop by double clicking on the app. It may take some time to start up.   

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_686")
```

#### Get workshop files
<input type="checkbox"> Download the files for this activity clicking here: https://github.com/fhdsl/containers-for-scientists-sandbox/archive/refs/heads/main.zip  
<input type="checkbox"> Put this file on your desktop so it is easily findable.  
<input type="checkbox"> Double click the zip file (or right click and choose "unzip" or "decompress" to unzip the file.  

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_693")
```

#### Activity 1: Pull and run a Docker image

**This activity is built so that you will encounter errors that this guide will work you through!**
Most common container stumbling block is understanding the idea that containers are isolated from your computer

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_700")
```

Which means: If your computer has files, software, or anything else. Your container by default does NOT have it unless you do something to get those files and software on there.

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_713")
```

We will encounter both an error of when we don't have a file we need AND an error of when we don't have a software package we need. We will guide you through what to do when these expected errors occur.

#### Step 1: Docker pull

We need to get the Docker **image** we want to use. We do this by "pulling" it.
Pulling is a term used in GitHub terminology too. It just means we're taking something from the internet and getting a copy locally (on our computer).

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_732")
```

<input type="checkbox"> Run this command in your Terminal or Command Prompt window:

```
docker pull cansav09/practice-image:1
```


```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_745")
```


##### Step 2: Check what images you have

<input type="checkbox"> To see what images we have we can run `docker image ls`.  
<input type="checkbox"> We should see `cansav09/practice-image:1` show up in the output.  

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_752")
```

##### Step 3: Run the docker image to start a container!

Remember Images can be shared easily but to actually use the stuff we need to turn our image into a container! To do that we will do `docker run`

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_760")
```

<input type="checkbox"> To launch the image for use, we can use `docker run` to turn it into a **container**.
```
docker run cansav09/practice-image:1
```

<input type="checkbox"> To see what containers we have running we have we can run:

```
docker ps
```

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_775")
```

Pat yourself on the back! You have a running container!

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_786")
```


What can we do with running containers?

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_791")
```

##### Step 4: Run a single command

<input type="checkbox"> To run stuff interactively from the command line we can do:  

```
docker exec -it <PUT_CONTAINER_ID_HERE> bash
```

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_803")
```

##### Step 5: Try calling a script

<input type="checkbox"> To run a script using the docker container we could just add reference to a script at the end:  

```
docker exec -it <PUT_CONTAINER_ID_HERE> bash run_analysis.sh
```

```{r, out.width = "100%", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g30a4ed49e59_0_812")
```

**BUT! You will find that this command won't work yet though, why?**

```
bash: run_analysis.sh: No such file or directory
```

**Question: Does our container have all of the same files that our computer has?**

This won't work because the file `run_analysis.sh` is not a file that our container has. Docker containers do not have all the files that our computer does; they only have the files we add to it.

#### Activity 1b: Add a volume to the image

Our container is separate from our computer so if we want to use a file we have to attach it using a "volume".

##### Step 1: Let’s add our containers-for-scientists-sandbox files

Let's point a volume to our workshop files so we have them on our container.

<details> </summary> Other ways we could get files on a container </summary>

If we want to run `run_analysis.sh` there's two ways we could get our files on there.
- We could download the files we need from online! We could `git clone` files from a repository or otherwise `wget` or `curl` files.
- We could add a `volume` of files that are local to our computer. For this we will use the You'll `-v` or volume option. We'll do this in the next step.

We can specify a particular file path on our computer or give it `$PWD` Then optionally we can give a `:` and a file path we'd like this to be stored on on the container. Otherwise it will be stored at the absolute top of the container.
</details>

<input type="checkbox"> Now we can run:  
```
docker run -v $pwd:/home cansav09/practice-image:1
```
If you have a windows machine you may have to run this variant instead. This version has a different `${}` around the `pwd` part.
```
docker run -v ${pwd}:/home cansav09/practice-image:1
```
#### Step 2: Retry calling the script

<input type="checkbox"> Now we can run the following command but we will have to run `docker ps` and get the container ID we need to put here.   
```
docker exec -it <REPLACE_WITH_CONTAINER_ID> bash /home/run_analysis.sh
```
Now we have a new error! What does this mean?

**Question: Does our container have all of the same software that our computer has?**

### Podman
#### Set up Podman
<input type="checkbox"> Go here to [install Podman](https://podman.io/docs/installation), following the instructions for your particular operating system.  
<input type="checkbox"> Open up your [command line](https://towardsdatascience.com/a-quick-guide-to-using-command-line-terminal-96815b97b955).  
<input type="checkbox"> Start up Podman by running the following:  
```
podman machine init
podman machine start
```

#### Get workshop files
<input type="checkbox"> Download the files for this activity clicking here: https://github.com/fhdsl/containers-for-scientists-sandbox/archive/refs/heads/main.zip  
<input type="checkbox"> Put this file on your desktop so it is easily findable.  
<input type="checkbox"> Double click the zip file (or right click and choose "unzip" or "decompress" to unzip the file.  

#### Activity 1: Pull and run an image

**This activity is built so that you will encounter errors that this guide will work you through!**
Most common container stumbling block is understanding the idea that containers are isolated from your computer

Which means: If your computer has files, software, or anything else. Your container by default does NOT have it unless you do something to get those files and software on there.

We will encounter both an error of when we don't have a file we need AND an error of when we don't have a software package we need. We will guide you through what to do when these expected errors occur.

##### Step 1: Podman pull

We need to get the Docker **image** we want to use. We do this by "pulling" it.

<input type="checkbox"> Run this command in your Terminal or Command Prompt window:

```
podman pull cansav09/practice-image:1
```

##### Step 2: Check what images you have

<input type="checkbox"> To see what images we have we can run `podman image ls`.  
<input type="checkbox"> You should see `cansav09/practice-image:1` show up in the output.  

##### Step 3: Run the docker image to start a container!

<input type="checkbox"> Now if want to launch the image for use, we can use `podman run` to turn it into a **container**.

```
podman run cansav09/practice-image:1
```

<input type="checkbox"> To see what containers we have running we have we can run:

```
podman ps
```

##### Step 4: Run a single command

<input type="checkbox"> To run stuff interactively from the command line we can do:

```
podman exec -it <PUT_CONTAINER_ID_HERE> bash
```

##### Step 5: Try calling a script

<input type="checkbox"> To run a script using the docker container we could just add reference to a script at the end:

```
podman exec -it <PUT_CONTAINER_ID_HERE> bash run_analysis.sh
```
**BUT! You will find that this command won't work yet though, why?**

This won't work because the file `run_analysis.sh` is not a file that our container has. Docker containers do not have all the files that our computer does; they only have the files we add to it.

```
bash: run_analysis.sh: No such file or directory
```

#### Activity 1b: Add a volume to the image

Our container is separate from our computer so if we want to use a file we have to attach it using a "volume".

##### Step 1: Let’s add our containers-for-scientists-sandbox files

Let's point a volume to our workshop files so we have them on our container.

<details> </summary> Other ways we could get files on a container </summary>
If we want to run `run_analysis.sh` there's two ways we could get our files on there.
- We could download the files we need from online! We could `git clone` files from a repository or otherwise `wget` or `curl` files.  
- We could add a `volume` of files that are local to our computer. For this we will use the You'll `-v` or volume option. We're going to do this in the next step.  
</details>

We will specify a particular file path on our computer or give it `$pwd` Then we can give a `:` and a file path we'd like this to be stored on on the container. Otherwise it will be stored at the absolute top of the container.

<input type="checkbox"> Now we can run:
```
podman run -v $pwd:/home cansav09/practice-image:1
```
If you have a windows machine you may have to run this variant instead. This version has a different `${}` around the `pwd` part.
```
podman run -v ${pwd}:/home cansav09/practice-image:1
```
##### Step 2: Retry calling the script

<input type="checkbox"> Now we can run:
```
podman exec -it <REPLACE_WITH_CONTAINER_ID> bash /home/run_analysis.sh
```

Now we have a new error! What does this mean?

**Question: Does our container have all of the same software that our computer has?**
