
# Advanced Container Techniques

```{r panel-setup, include = FALSE}
xaringanExtra::use_panelset()
xaringanExtra::style_panelset_tabs(font_family = "inherit")
```


## GitHub Actions to manage your containers 

GitHub Actions pairs really well with images [We encourage you to take a look at our course on GitHub Actions for science!](https://hutchdatascience.org/GitHub_Automation_for_Scientists/)

GitHub Actions can use Docker containers really seamlessly to automate things for you. 
GitHub Actions can also be a way for you to manage your Docker images and Dockerfiles. 

See an example of one we actually maintain here: https://github.com/jhudsl/ottr-docker 


## Personal Access Tokens 

To give GitHub Actions or other items access to your Dockerhub you will need to use Personal access Tokens. 
As with most Personal Access Tokens, be very careful to not share these or post them anywhere publicly!

```{r, out.width = "100%", fig.alt = "Personal access tokens can be used to help manage and integrate your docker images from elsewhere.", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g2effc5b673e_0_647") 
```

## Docker compose 

In this course, we introduced you to docker run. But what if you have a ton of arguments to specify and things that you can't do from a single command? Enter docker compose. This allows you to get very fancy with your docker containers and specify a bunch of things in your yaml file. 

```{r, out.width = "100%", fig.alt = "What if I have a lot of arguments I need to specify? Docker compose. Specify options in a yml file", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g2effc5b673e_0_655") 
```

## What's Kubernetes

[Kubernetes is a way you can use a bunch of containers at once](https://kubernetes.io/docs/concepts/overview/). It's mostly not needed for scientific analyses but now you know about it. 

```{r, out.width = "100%", fig.alt = "What if I have a lot of containers I need to coordinate? Kubernetes", echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1T5Lfei2UVou9b0qaUCrWXmkcIwAao-UcN4pHMPEE4CY/edit#slide=id.g2effc5b673e_0_660") 
```
